---
title: "penguin_analysis"
format: html
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Week 1

# Loading libraries and code files.

```{r}
library(tidyverse)
library(palmerpenguins)
library(janitor)
library(here)

#source function says to R i want you to find this file and use the stuff within this file to run the code laid out below.
source(here("functions","cleaning_penguins.r"))
```

# Preserving Raw Data

```{r}
write.csv(penguins_raw, here("data","penguins_raw.csv"))

#loading raw data
penguins_raw<-read.csv(here("data","penguins_raw.csv"))
```

# Cleaning raw data and generating clean data set.
```{r}
#Using functions from functions script (cleaning.R)
cleaning_penguin_columns <- function(raw_data){
  print("cleaned names, removed comments, empty rows and columns and delta")
  raw_data %>% 
    clean_names() %>% 
    shorten_species() %>% 
      remove_empty(c("rows", "cols")) %>% 
      select(-starts_with("delta")) %>% 
      select(-comments)}
      
#Runs code
colnames(penguins_raw)
penguins_clean<-cleaning_penguin_columns(penguins_raw)
colnames(penguins_clean)

#writes new csv for clean penguins data. 
write.csv(penguins_clean, here("data","penguins_clean.csv"))
```

# Removing NA values:
```{r}
body_mass <- penguins_clean %>%
    select(species, body_mass_g) %>%
    remove_NA()

print(paste("Number of rows:", nrow(body_mass)))
print(paste("Number of missing values:", sum(is.na(body_mass))))
```

# Filtering by Species
```{r}
adelie_data <- penguins_clean %>%
  filter(species == "Adelie")

head(adelie_data)
```

# Combining the above 2:
```{r}

adelie_body_mass <- penguins_clean %>%
  filter(species == "Adelie") %>%
  select(species, body_mass_g) %>%
  remove_NA()

head(adelie_body_mass)
```

# Renv:

should all be typed in console. Here, am using \# to ensure code does not accidentally run. renv::init initialises renv. renv::snapshot creates renv.lock file, allowing us to keep a record of all packages installed for this analysis. renv::diagnostics allows us to run a summary of what packages have been installed by renv. renv::restore() allows us or someone else to restore the snapshot and ensure the same libraries are installed. renv::status allows us to check if renv is active. Anything with a :: should be typed in the console according to LF.

# Part2 
```{r}
here::here()

#Tells where to find file on my machine. penguins_clean \<- read.csv(here("data", "penguins_clean.csv"))

#codes for boxplot flipper_boxplot_old \<- ggplot( data=penguins_clean, aes(x=species, y=flipper_length_mm)) + geom_boxplot()

#codes boxplot flipper_boxplot_old

#new dataset to get rid of NA vals only in flipper column (only has flipper length and species). penguins_flippers \<- select(penguins_clean, c("species","flipper_length_mm")) %\>% drop_na() colnames(penguins_flippers) summary(penguins_flippers)

#NOTE- she has just taken code from above, called boxplot penguins_flippers and deleted old code. I have kept to show what I have done. species_colours \<- c("Adelie"="darkorange", "Chinstrap"="purple", "Gentoo"="cyan4")

flipper_boxplot \<- ggplot( data=penguins_flippers, aes(x=species, y=flipper_length_mm))+ geom_boxplot(), width=0.7, show.legend=FALSE )

flipper_boxplot

species_colours \<- c("Adelie"="darkorange", "Chinstrap"="purple", "Gentoo"="cyan4")
